Clicks and Cliques

### 2018-10-22 papersoul.Rnw
include guest editor's comments (ecxept sampling weights)

\begin{table}[h]
\centering
\caption{Sources of information and variables used in the analysis \label{des.var}}
\begin{tabular}{|c|l|p{10cm}|}\hline
 & \textbf{Variable Group}&\textbf{Description}\\ \hline
 \multicolumn{1}{ |c| }{\multirow{4}{*}[-12pt]{\rotatebox[origin=c]{90}{Knight survey data}}}
 & Original variables & There are 118 original variables, most of them consist of ordinal variables with a scale from 1 to 5. All these variables ask for the opinion about some aspect of community life.\\ \cline{2-3}
& Demographic Variables &There are 30 demographic variables, these variables have information about, age, race, gender, etc.\\ \cline{2-3}
 & Recoded Variables&Some of questionnaire variables has been recoded in a scale from 1 to 3. \\ \cline{2-3}
 & Constructed variables & Using recoded variables, some summary variables have been made. These new variables are the base for the Gallup analysis of this data and represent the different sociological dimensions to be study. Some of these variables measure, economy level, social offerings, openness, safety, etc\\ \hline
 \multicolumn{1}{ |c| }{\multirow{7}{*}[-30pt]{\rotatebox[origin=c]{90}{Census Bureau data}}}
& Age-Sex &  Total of people on each age group by gender. Proportions of female and the average age for each gender were computed \\ \cline{2-3}
 & Education & Total of people between 25-64 years old, divided by education level attained and working status. Unemployment rate and the proportion of people within each education level were computed. \\ \cline{2-3}
 & Income &People in each level of income. We computed average income, income per-capita and Gini coefficient\\ \cline{2-3}
 & Owner &Information relative to the total of people living in houses which are renters or are owners, and also how long they are living in that place. We computed average years that people have lived in the same house separately for renters and owners.\\ \cline{2-3}
 & Race &People totals in each race, mainly white, black and Other.\\ \cline{2-3}
 & Workers & Information relative to the household size and the workers per household.\\ \cline{2-3}
 & Year of entry &Total of people which has born outside the USA, and how long they lived in the community.\\ \hline
\end{tabular}
\end{table}

The Knight survey data contains missing information. The proportion of missing values was computed for each observation (by cases) and for each variable (by variables). Table \ref{tab1} shows the quantiles of the proportion of missing values computed by case and computed by variable. Focusing on the proportions of missing values by case, there is 19\% of missing record for the most complete response. This may be explained by the fact that 3 years of data was combined although some variables only appeared in 1 or 2 years. On the other hand, focusing on the proportion of missing values by variables, most of the variables show a reasonable proportion of missing values (the median is 12\%) while there are a few variables with really high proportions of missing values.

@@

Supervised and unsupervised learning tools were used in this paper to answer the main questions presented in Section \ref{intro}. To present the results of statistical learning methods, different visualization devices were used to perform an exploratory analysis. In this section the main tools for graphical and statistical analysis are described.

@@

Product plot is a very flexible framework, more than 20 visualizations previously described can be included in it \citep{wickham2011product}, like bar charts, spine plots \citep{hummel1996linked} and mosaic plots \citep{hartigan1981mosaics,friendly1994mosaic}. This framework is designed for study relations among categorical variables. However, it is easily adapted when the variables are continuous simply by binning variables.

@@

Cluster analysis is an unsupervised technique and two ways to find clusters are used across this paper. Hierarchical agglomerative cluster is used to find clusters at the community level. This method starts with each point being its own cluster and they are recursively merged. Then, the ''merging history'' forms a hierarchical tree diagram called a dendrogram, which is a tool used to define the number of clusters (see Sec. 12.3.1 of \cite{izenman2008modern} for more details).

Also, K-means clustering was used to find clusters at the individual level. The K-means algorithm, proposed by \cite{macqueen1967some}, starts with all observations assigned to $K$ initial clusters and cluster means are computed. Then in an iterative fashion observations are reassigned to minimize the within sums of squares of each cluster until a criterion of convergence has been met (see Sec. 12.4.1 of \cite{izenman2008modern} for more details).

@@

\subsubsection{Random Forest}
Random forest is a supervised learning technique proposed by \cite{breiman2001random}, based on tree predictors with extra randomness added in its construction. A tree predictor is a classifier which recursively partition the feature space looking for regions where the response variable is homogeneous (pure nodes). In order to build a forest, bootstrap samples from the data are selected and for each sample a tree classifier is grown. However, the constructed tree in each bootstrap sample only uses a random selection of variables to find the next feature partition. Then two sources of randomness are used, random selection of cases in the bootstrap sample and random selection of variables. Then the forest classifier is an ensemble of randomized trees, the forest prediction is determined with majority vote across trees (see Section 14.4 of \cite{izenman2008modern} for more details).

@@

p <- ggmap(mapp, extent = "device")
p + geom_point(data = aux.mer, mapping = aes(lon, lat), size = 6*aux.mer$cate, alpha = .4) + xlab('') + ylab('') + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
@
  \caption{Location of the 26 communities analyzed, size of dot represent population. Only two communities on the west coast. \label{map}}
\end{figure}

@@

@
\caption{Proportion of people by age category, horizontal axis is the midpoint age for each group. The line color and point shape represents the gender. Each facet is a community, sorted from top to bottom according to proportion of people between 18 and 21 years old. Similar age distribution across gender in all communities. Only State College and Grand Folks shows a peak on 18-21 years old. \label{age}}
\end{figure}

@@

census.data %>% select( c(colnames(census.data)[var.cl2],'cl.2' ) ) %>% 
  group_by(cl.2) %>%
  summarise_all(funs(mean)) %>%
  setNames(nn) %>% ungroup() %>%
  mutate(cl.2 = factor(cl.2)) %>%
  ggparcoord(columns = 2:10, scale = 'std', 
             groupColumn = 'cl.2', showPoints = T, 
             order = c(3, 4, 6, 10, 2, 5, 9, 7, 8))  + 
  theme(axis.text.x = element_text(angle = 90, size = I(15))) + 
  xlab('') + ylab('') + 
  scale_color_brewer(type = 'qual', palette = 2, name = 'Group')
@
\caption{Parallel coordinate plot for community cluster means \label{cluster.cen}.  }
\end{figure}

@@

To characterize the cluster a parallel coordinate plot \citep{inselberg1985plane} of the cluster means was used. Figure \ref{cluster.cen} shows the variables used in the horizontal axis and the variable mean of each cluster on the vertical axis. Group 1 are communities with older people, low education level and biggest home owner rate. Group 2 is formed by communities with large proportion of African-Americans and unequal income distribution. Bradenton and Palm Beach (entire group 3) are communities with older people, not working and not African-American. Group 4 is formed by four communities: Charlotte, Grand Forks, Long Beach and State College, all relatively younger people, with high education levels and non equitable income distribution.

@@

#,main='Mean Value Selected Variables Used in the K-means Cluster'
qplot(data = aux2, var2, value, color = Clusters,shape = Clusters,xlab = '', ylab = '',size = I(3)) +geom_line(size = I(1)) + facet_wrap(~YEAR,nrow = 3) + theme(axis.text.x = element_text(angle = 90, size = I(15)), legend.position = 'bottom' ) + scale_x_discrete(breaks = 1:31, labels = lab) + scale_color_brewer(type='qual', palette='Paired')
@
\caption{Cluster mean points. Mean values for all the variables used to create the clusters in every year. \label{para}}
\end{figure}

@@

@
\caption{Relation between community attachment and k-means cluster solution.\label{clus1}}
\end{figure}

@@

Table \ref{move} shows that among the attached people 19\% would like to move if they had the chance to do it and this proportion increases to 70\% among not attached people. On the other hand, 29 \% of the not attached people would like to stay in the same neighborhood while among attached people this proportion reaches 80\%. This suggests a positive association among the attached group and a longer stay in one community.

@@

@
\caption{Dendrogam of communities base on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}

@@

\begin{acknowledgements}
Professor Di Cook helped with the analysis and reviewing the paper. Xiaoque Cheng's mergeGui was used to create more amenable initial data. Israel Almodovar, Vianey Leos and Ricardo Martinez help us a lot proofreading our paper for final corrections.
\end{acknowledgements}

>>>>

\begin{table}[h]
\centering
\caption{Sources of information and variables used in the analysis. \label{des.var}}
\begin{tabular}{|c|l|p{10cm}|}\hline
 & \textbf{Variable Group}&\textbf{Description}\\ \hline
 \multicolumn{1}{ |c| }{\multirow{4}{*}[-12pt]{\rotatebox[origin=c]{90}{Knight survey data}}}
 & Original variables & There are 118 original variables, most of them consist of ordinal variables with a scale from 1 to 5. All these variables ask for the opinion about some aspect of community life.\\ \cline{2-3}
& Demographic variables &There are 30 demographic variables, these variables have information about, age, race, gender, etc.\\ \cline{2-3}
 & Recoded variables & Some of questionnaire variables has been recoded in a scale from 1 to 3. \\ \cline{2-3}
 & Constructed variables & Using recoded variables, some summary variables have been made. These new variables are the base for the Gallup analysis of this data and represent the different sociological dimensions to be study. Some of these variables measure, economy level, social offerings, openness, safety, etc\\ \hline
 \multicolumn{1}{ |c| }{\multirow{7}{*}[-30pt]{\rotatebox[origin=c]{90}{Census Bureau data}}}
& Age-Sex &  Total of people on each age group by gender. Proportions of female and the average age for each gender were computed \\ \cline{2-3}
 & Education & Total of people between 25-64 years old, divided by education level attained and working status. Unemployment rate and the proportion of people within each education level were computed. \\ \cline{2-3}
 & Income &People in each level of income. We computed average income, income per-capita and Gini coefficient\\ \cline{2-3}
 & Owner &Information relative to the total of people living in houses which are renters or are owners, and also how long they are living in that place. We computed average years that people have lived in the same house separately for renters and owners.\\ \cline{2-3}
 & Race &People totals in each race, mainly white, black and Other.\\ \cline{2-3}
 & Workers & Information relative to the household size and the workers per household.\\ \cline{2-3}
 & Year of entry &Total of people which has born outside the USA, and how long they lived in the community.\\ \hline
\end{tabular}
\end{table}

The Knight survey data contains missing information. The proportion of missing values was computed for each observation (by variables) and for each variable (by cases). Table \ref{tab1} shows the quantiles of the proportion of missing values computed by case and computed by variable. Focusing on the proportions of missing values by variables, there is 19\% of missing record for the most complete response. This may be explained by the fact that 3 years of data was combined although some variables only appeared in 1 or 2 years. On the other hand, focusing on the proportion of missing values by casess, most of the variables show a reasonable proportion of missing values (the median is 12\%) while there are a few variables with really high proportions of missing cases.

@@

Supervised and unsupervised learning tools were used in this paper to answer the main questions presented in Section \ref{intro}. To present the results of statistical learning methods, different visualizations were used to perform an exploratory analysis. In this section the main tools for graphical and statistical analysis are described.

@@

Product plot is a very flexible framework \citep{wickham2011product}, which incorporate displays like bar charts, spine plots \citep{hummel1996linked} and mosaic plots \citep{hartigan1981mosaics,friendly1994mosaic}. This framework is designed for study relations among categorical variables. However, it is easily adapted when the variables are continuous simply by binning variables.

@@

Cluster analysis is an unsupervised technique and two ways to find clusters are used across this paper. Hierarchical agglomerative cluster is used to find clusters at the community level. This clustering algorithm starts with each point being its own cluster and is recursively merged. Then, the ``merging history'' forms a hierarchical tree diagram called a dendrogram, which is a tool used to define the number of clusters (see Sec. 12.3.1 of \cite{izenman2008modern} for more details). Additionally, K-means clustering was used to find clusters at the individual level. The K-means algorithm, proposed by \cite{macqueen1967some}, starts with all observations assigned to $K$ initial clusters and cluster means are computed. Then in an iterative fashion observations are reassigned to minimize the within sums of squares of each cluster until a criterion of convergence has been met (see Sec. 12.4.1 of \cite{izenman2008modern} for more details).

@@

\subsubsection{Random forest}
Random forest is a supervised learning technique proposed by \cite{breiman2001random}, based on combining results from many tree predictors computed on bootstrap samples of the data. A tree predictor is a classifier which recursively partition the feature space looking for regions where the response variable is homogeneous (pure nodes). In order to build a forest, bootstrap samples from the data are selected and for each sample a tree classifier is grown. However, the constructed tree in each bootstrap sample only uses a random selection of variables to find the next feature partition. Then two sources of randomness are used, random selection of cases in the bootstrap sample and random selection of variables. Then the forest classifier is an ensemble of randomized trees, the forest prediction is determined with majority vote across trees (see Section 14.4 of \cite{izenman2008modern} for more details).

@@

p <- ggmap(mapp, extent = "device")
p + geom_point(data = aux.mer, mapping = aes(lon, lat), size = 6*aux.mer$cate, alpha = .4) + xlab('') + ylab('') + theme(axis.text.x = element_blank(), axis.text.y = element_blank())
@
  \caption{Location of the 26 communities analyzed, size of dot represent population. Only two communities on the west coast are included. \label{map}}
\end{figure}

@@

@
\caption{Proportion of people by age category. The line color and point shape represents the gender. Each facet is a community, sorted from top to bottom according to proportion of people between 18 and 21 years old. Only State College and Grand Folks shows a peak on 18-21 years old. \label{age}}
\end{figure}

@@

census.data %>% select( c(colnames(census.data)[var.cl2],'cl.2' ) ) %>% 
  group_by(cl.2) %>%
  summarise_all(funs(mean)) %>%
  setNames(nn) %>% ungroup() %>%
  mutate(cl.2 = factor(cl.2)) %>%
  ggparcoord(columns = 2:10, scale = 'std', 
             groupColumn = 'cl.2', showPoints = T, 
             order = c(3, 4, 6, 10, 2, 5, 9, 7, 8))  + 
  theme(axis.text.x = element_text(angle = 90, size = I(15))) + 
  xlab('') + ylab('') + 
  scale_color_brewer(type = 'qual', palette = 2, name = 'Group')
@
\caption{Parallel coordinate plot for community cluster means. \label{cluster.cen} }
\end{figure}

@@

A parallel coordinate plot \citep{inselberg1985plane} is used to compare the cluster means for the variables used in the analysis. Figure \ref{cluster.cen} shows the variables used in the horizontal axis and the variable mean of each cluster on the vertical axis. Group 1 are communities with older people, low education level and biggest home owner rate. Group 2 is formed by communities with large proportion of African-Americans and unequal income distribution. Bradenton and Palm Beach (entire group 3) are communities with older people, not working and not African-American. Group 4 is formed by four communities: Charlotte, Grand Forks, Long Beach and State College, all relatively younger people, with high education levels and non equitable income distribution.

@@

#,main='Mean Value Selected Variables Used in the K-means Cluster'
qplot(data = aux2, var2, value, color = Clusters,shape = Clusters,xlab = '', ylab = '',size = I(3)) +geom_line(size = I(.5)) + facet_wrap(~YEAR,nrow = 3) + theme(axis.text.x = element_text(angle = 90, size = I(15)), legend.position = 'bottom' ) + scale_x_discrete(breaks = 1:31, labels = lab) + scale_color_brewer(type='qual', palette='Paired')
@
\caption{Cluster mean points. Mean values for all the variables used to create the clusters in every year. \label{para}}
\end{figure}

@@

@
\caption{Relation between community attachment and k-means cluster solution. Cluster 1 shows the higest attachment proportion while cluster 2 the lowest. Attachment distribution is stable within each grupo amont years. \label{clus1}}
\end{figure}

@@

Table \ref{move} shows that among the attached people 19\% would like to move if they had the chance to do it and this proportion increases to 70\% among not attached people. On the other hand, 29\% of the not attached people would like to stay in the same neighborhood while among attached people this proportion reaches 80\%. This suggests a positive association among the attached group and a longer stay in one community.

@@

@
\caption{Dendrogam of communities based on importance measures. Communities close to each other have similar structure of key drivers of attachment. \label{dendro}}
\end{figure}

@@

\begin{acknowledgements}
Professor Di Cook helped with the analysis and reviewing the paper. Xiaoyue Cheng's mergeGui was used to create more amenable initial data. Israel Almodovar, Vianey Leos and Ricardo Martinez helped us with proofreading our paper for final corrections.
\end{acknowledgements}
